PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
```

@@

# Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.edgi",
                        "Age_uncleaned",
                        "EduYrs")

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.edgi = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.edgi, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      EduYrs = demographics.school_education_school_years.txt, # Years in education
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r nbr copy gad}
if(NBR == TRUE){
gad.raw.id <- glad.gad.raw.id
}
```

@@

```{r edgi gad cleaning}
exclude_cols <- c("Sample",
                  "ID")

@@

# Inspect dimensions
dim(edgi.gad.raw.id)
# Inspect colnames
colnames(edgi.gad.raw.id)
#Differences
dim(edgi.gad.raw)[1]-dim(edgi.gad.raw.id)[1]
```

```{r nbr copy gad}
if(NBR == TRUE){
  gad.raw.id <- edgi.gad.raw.id
}
```

@@

# Calulate total score
gad.scored_items_prepan <- scoreItems(keys = gad.items_key_prepan,
                                        items = glad.gad.raw.id[gad.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
GADraw.id$gad.sum_score_prepan_raw <- gad.scored_items_prepan$scores

# Look at values
GADraw.id %>%
  descr(gad.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r gad prepan missingness}
glad.gad.raw.id <- gad.raw.id %>%
  mutate(
    na_per_person_gad_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[gad.items_prepan])
      )
  )

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = TRUE
EDGI = FALSE
NBR = FALSE
RAMP = FALSE
```

@@

# Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID),
           startDate.edgi = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.edgi, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r glad copy gad}
if(GLAD == TRUE){
gad.raw.id <- glad.gad.raw.id
}
```

@@

```{r edgi gad cleaning}
if(EDGI == TRUE){
exclude_cols <- c("Sample",
                  "ID")

@@

# Inspect dimensions
dim(edgi.gad.raw.id)
# Inspect colnames
colnames(edgi.gad.raw.id)
#Differences
dim(edgi.gad.raw)[1]-dim(edgi.gad.raw.id)[1]
}
```

```{r nbr copy gad}
if(EDGI == TRUE){
  gad.raw.id <- edgi.gad.raw.id
}
```

@@

# Calulate total score
gad.scored_items_prepan <- scoreItems(keys = gad.items_key_prepan,
                                        items = gad.raw.id[gad.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
gad.raw.id$gad.sum_score_prepan_raw <- gad.scored_items_prepan$scores

# Look at values
gad.raw.id %>%
  descr(gad.sum_score_prepan_raw,
        stats = "common"
        )
```

@@

prepandemic
```{r gad prepan missingness}
gad.raw.id <- gad.raw.id %>%
  mutate(
    na_per_person_gad_prepan =         # variable for NA per person
      rowSums(                       
        is.na(.[gad.items_prepan])
      )
  )
